Computing with an Angry Poker Player

### 2014-09-25 draft0.tex
fixed figure reference

One can get another trivial upper bound for $T(n)$ by looking at the maximum value of the potential function. Remember that the board will only contain chips in the areas denotes by locations in $[-\sqrt{n}, \sqrt{n}] \times [-\sqrt{n}, \sqrt{n}]$. So we know that when the board is stable at some $f'$, 
\[ \Phi(f') \leq \sum_{x,y} f'(x,y)(|x| + |y|)^2 \leq 4n*3*4n = O(n^2) \]
And since $\Phi(f) = 0$ and we know that $\Phi$ increases by at least $4$ in each global earthquake, 
\[ T(n) = O(n^2) \]
We can get tighter bounds by looking at the number of earthquakes more carefully. Lets let $T(x,y)$ be the number of local earthquakes needed to hit location $(x,y)$ for the board to stabilize. Suppose $(x,y) \neq (0,0)$, then the number of earthquakes required to hit location $(x,y)$ will be the total number of chips entering location $(x,y)$ divided by 4 (taking floors ofcourse). In particular,
\[ T(x,y) = \left\{ \begin{array}{cc} \lfloor\frac{T(x+1, y)}{4} \rfloor + \lfloor \frac{T(x-1, y)}{4} \rfloor + \lfloor \frac{T(x, y-1)}{4} \rfloor + \lfloor \frac{T(x, y+1)}{4} \rfloor & (x,y) \neq (0,0) \\
 \lfloor \frac{n}{4} \rfloor  + \lfloor\frac{T(x+1, y)}{4} \rfloor + \lfloor \frac{T(x-1, y)}{4} \rfloor + \lfloor \frac{T(x, y-1)}{4} \rfloor + \lfloor \frac{T(x, y+1)}{4} \rfloor & (x,y) = (0,0) \end{array} \right. \]
So solving for $T$ is a solution to a discrete version of the Laplacian. We were able to experiment different scenarious to understand the growth of $T(n)$. 
\begin{figure}[!ht]
\centering
\subfigure[Global earthquakes until stable. Horizontal axis is $n$ and vertical axis is the number of global earthquakes]{\includegraphics[width=0.4\linewidth]{TimeUntilStabilizeWithOnePeak}\label{fig:case1}} \qquad
\subfigure[Global earthquakes until stable. Horizontal axis is $n$ and vertical axis is the number of global earthquakes divided by $n$]{\includegraphics[width=0.4\linewidth]{Timeuntilstabledividedbyn}\label{fig:case2}}
\caption{Number of global earthquakes until stability}
\label{fig:growthofT}
\end{figure}
By looking at Figure~\ref{fig:growthofT} seems that $T(n) = \Theta(n\log n)$. Our strategy to show a lower bound on $T(n)$ by giving a lower bound on $T'(n)$ seems to be the correct approach, since by experimental evidence, we believe that $T'(n) = \Theta(n\log n)$.

\begin{figure}[!ht]
\centering
\subfigure[Local earthquakes of center until stable. Horizontal axis is $n$ and vertical axis is the number of local earthquakes]{\includegraphics[width=0.4\linewidth]{centerearthquakes}\label{fig:case1}} \qquad
\subfigure[Local earthquakes of center until stable. Horizontal axis is $n$ and vertical axis is the number of local earthquakes divided by $n$]{\includegraphics[width=0.4\linewidth]{centerearthquakesdividedbyn}\label{fig:case2}}
\caption{Number of local earthquakes in center until stability}
\label{fig:growthofT}
\end{figure}

>>>>

One can get another trivial upper bound for $T(n)$ by looking at the maximum value of the potential function. Remember that the board will only contain chips in the areas denotes by locations in $[-\sqrt{n}, \sqrt{n}] \times [-\sqrt{n}, \sqrt{n}]$. So we know that when the board is stable at some $f'$, 
\[ \Phi(f') \leq \sum_{x,y} f'(x,y)(|x| + |y|)^2 \leq 4n*3*4n = O(n^2) \]
And since $\Phi(f) = 0$ and we know that $\Phi$ increases by at least $4$ in each global earthquake, 
\[ T(n) = O(n^2) \]
We can get tighter bounds by looking at the number of earthquakes more carefully. Lets let $T(x,y)$ be the number of local earthquakes needed to hit location $(x,y)$ for the board to stabilize. Suppose $(x,y) \neq (0,0)$, then the number of earthquakes required to hit location $(x,y)$ will be the total number of chips entering location $(x,y)$ divided by 4 (taking floors ofcourse). In particular,
\[ T(x,y) = \left\{ \begin{array}{cc} \lfloor\frac{T(x+1, y)}{4} \rfloor + \lfloor \frac{T(x-1, y)}{4} \rfloor + \lfloor \frac{T(x, y-1)}{4} \rfloor + \lfloor \frac{T(x, y+1)}{4} \rfloor & (x,y) \neq (0,0) \\
 \lfloor \frac{n}{4} \rfloor  + \lfloor\frac{T(x+1, y)}{4} \rfloor + \lfloor \frac{T(x-1, y)}{4} \rfloor + \lfloor \frac{T(x, y-1)}{4} \rfloor + \lfloor \frac{T(x, y+1)}{4} \rfloor & (x,y) = (0,0) \end{array} \right. \]
So solving for $T$ is a solution to a discrete version of the Laplacian. We were able to experiment different scenarious to understand the growth of $T(n)$. 
\begin{figure}[!ht]
\centering
\subfigure[Global earthquakes until stable. Horizontal axis is $n$ and vertical axis is the number of global earthquakes]{\includegraphics[width=0.4\linewidth]{TimeUntilStabilizeWithOnePeak}\label{fig:case1}} \qquad
\subfigure[Global earthquakes until stable. Horizontal axis is $n$ and vertical axis is the number of global earthquakes divided by $n$]{\includegraphics[width=0.4\linewidth]{Timeuntilstabledividedbyn}\label{fig:case2}}
\caption{Number of global earthquakes until stability}
\label{fig:growthofT}
\end{figure}
By looking at Figure~\ref{fig:growthofT} seems that $T(n) = \Theta(n\log n)$. Our strategy to show a lower bound on $T(n)$ by giving a lower bound on $T'(n)$ seems to be the correct approach, since by experimental evidence (see Figure~\ref{fig:growthofT'}, we believe that $T'(n) = \Theta(n\log n)$.

\begin{figure}[!ht]
\centering
\subfigure[Local earthquakes of center until stable. Horizontal axis is $n$ and vertical axis is the number of local earthquakes]{\includegraphics[width=0.4\linewidth]{centerearthquakes}\label{fig:case1}} \qquad
\subfigure[Local earthquakes of center until stable. Horizontal axis is $n$ and vertical axis is the number of local earthquakes divided by $n$]{\includegraphics[width=0.4\linewidth]{centerearthquakesdividedbyn}\label{fig:case2}}
\caption{Number of local earthquakes in center until stability}
\label{fig:growthofT'}
\end{figure}
